\documentclass[11pt,oneside]{amsart}
\input{preamble}

\title{Problem ideas}

\begin{document}
  \maketitle

  \begin{problem}
    Let $\bw=(1,1,\dots,1)$. Let $\bv=(v_1,\dots,v_n)$. Prove that $\mathbf P_\bw\bv=(x,x,\dots,x)$ where $x$ is the average of $v_1,\dots,v_n$.
  \end{problem}

  \begin{problem}
    Give a better proof of the theorem that ($\bv,\bw\neq 0$) $\bv\|\bw\iff \bv=\lambda\bw$ for some $\lambda$, than what is in the textbook. (I have it in mind, just too lazy to write it out here for the moment.)
  \end{problem}

  \begin{problem}
    A conceptual proof of the sine angle formula for cross product??? (I haven't found one yet)
  \end{problem}

  \begin{problem}
    The graph of the function $f(x,y)=\frac35x+\frac45y$ is a plane. What is the normal vector to this plane?
  \end{problem}

  \begin{problem}
    (For exam) Make a vector-valued function whose trajectory is a line but for which the speed is not constant.
  \end{problem}

  \begin{problem}\leavevmode
    \begin{enumerate}[(a)]
      \item (7 points) Suppose $f(x,y)$ is a function and $c_1,c_2$ are two \textbf{different} real numbers. Prove by contradiction that it is impossible for the same point $(x_0,y_0)$ to be in both the level set $L_{c_1}$ and the level set $L_{c_2}$.

      Hint: Start with the following sentence: ``Suppose for the sake of contradiction that $(x_0,y_0)\in L_{c_1}$ and $(x_0,y_0)\in L_{c_2}$.''
      \vfill
      \vfill
      \item (3 points) Explain why part (a) proves that level curves of any function $f(x,y)$ cannot intersect each other.
      \vfill
    \end{enumerate}
  \end{problem}

  \begin{problem}
    Figure out various angles in a regular tetrahedron inscribed in a cube.
  \end{problem}

  \begin{problem}
    Derive the equation of a cardioid based on a circle rolling around another circle.
  \end{problem}

  \begin{problem}
    If $\br(t)$ is an integral curve for $\BF$, and $C$ is a curve parametrized by $\br(t)$ for $a\leq t\leq b$, prove that $\int_C\BF\cdot d\br\geq 0$.
  \end{problem}

  \begin{problem}
    The flux of the curl of $F$ through a closed surface is 0.
  \end{problem}

  \begin{problem}
    Parametrize the surface $z=\sqrt{1-x^2-y^2}$ as a vector-valued function $\br(u,v)$.
  \end{problem}

  \begin{problem}
    Show that $d\mathbf S=\mathbf n\cdot dS$.
  \end{problem}

  \begin{problem}
    Let $\br(u,v)=$(TODO). Determine a formula for the direction of $\br_u\times \br_v$, i.e.\ what is $\dir(\br_u\times\br_v)$? Express your answer in terms of $u$ and $v$.
  \end{problem}
  \begin{problem}
    Prove that (TODO) also gives you the area of the enclosed region, similar to (13.5.4).
  \end{problem}
\end{document}
