\documentclass[11pt,oneside]{amsart}
\input{../preamble}

\title{MATH2202 Spring 2024\\
Discussion 3}
\date{February 6, 2024}

\theoremstyle{definition}
\newtheorem{question}{Question}

\begin{document}
  \maketitle

  \begin{problem}
    The TA will demonstrate parts (a) and (b) of this question. (The TA may also want to give some hints for (c).)
    \begin{enumerate}[(a)]
        \item (Relevant to a homework question) Write $\frac{2-t}{1-t}$ in the form $f(\phi(t))$ where $f$ is a real-valued function and $\phi$ is defined by $\phi(t)=\frac1{1-t}$.
        \item Suppose that the speed $\|\br'(t)\|$ is the zero function. Prove that $\br(t)$ is constant.
        \item Suppose that the speed $\|\br'(t)\|$ is constant but nonzero. Does it follow that $\br(t)$ is linear?
    \end{enumerate}
  \end{problem}

  \begin{problem}
      \leavevmode\begin{enumerate}[(a)]
          \item Find a parametrization of the line through $(3,1,5)$ and $(5,1,3)$.
          \item What is an equation of the plane through $(1,0,0)$, $(0,1,0)$, and $(0,0,1)$?
          \item What is an equation of the plane through $(2,0,0)$, $(0,2,0)$, and $(0,0,2)$? Try to do less work than doing the problem all over again, by taking advantage of the answer to part (b).
          \item The part of the plane from (b) with all three coordinates non-negative forms a triangle. What is the area of this triangle? What about that of part (c)?
      \end{enumerate}
  \end{problem}

  \begin{problem}
    \leavevmode\begin{enumerate}[(a)]
        \item If $\br(t)$ is the position of a particle at time $t$, how do you interpret the function $\|\br(t)\|$? (Notice: no derivative is being taken, so this isn't speed.) For example, is this a vector-valued function or scalar-valued function?
        \item Plot $\br(t),\br'(t)$, and $\br''(t)$ for the helix $\br(t)=(\cos(t),\sin(t),t)$.
        \item Plot $\br(t)$ and $\br'(t)$ for the reparametrized helix $(\cos(t^2),\sin(t^2),t^2)$. Don't bother with $\br''(t)$, it'll be messy.
    \end{enumerate}
  \end{problem}

\subsection*{Challenge problem}
Pick a tilted circle in space (tilted meaning not parallel to any of the standard axes or planes) and figure out a parametrization for it. Argue that the trajectory of your parametrization traces out a circle. (Hint: What if you wrote a non-tiled circle in terms of $\bi,\bj$, then replaced $\bi$ and $\bj$ with another pair of perpendicular vectors of your choosing?)

\end{document}
