\documentclass[11pt,oneside]{amsart}
\input{../preamble}

\title{MATH2202 Spring 2024\\
Discussion 2}
\date{January 30, 2024}

\theoremstyle{definition}
\newtheorem{question}{Question}

\begin{document}
  \maketitle

  \begin{problem}
    The TA will demonstrate parts (a) and (b) of this question.
    \begin{enumerate}[(a)]
        \item Find a normal vector to the plane with equation $2(x-y+z)=3x+4y+5$.
        \item Find the plane with normal vector $(1,-3,-3)$ and passing through the point $(2,0,-1)$.
        \item Fully generalize part (b): Find the plane with normal vector $(a,b,c)$ and passing through the point $(x_0,y_0,z_0)$.
    \end{enumerate}
  \end{problem}

  \begin{problem}
      \leavevmode\begin{enumerate}[(a)]
          \item Let's calculate some cross products: let $\bv=(1,2,3)$ and $\bw=(-3,-2,-1)$. Calculate $\bv+\bw$, $\bv-\bw$, $\bv\times\bw$, and finally $(\bv+\bw)\times(\bv-\bw)$. What do you notice?
          \item Generalize the above: Prove that $(\bv+\bw)\times(\bv-\bw)=-2(\bv\times\bw)$ for all $\bv,\bw$. There is a nicer proof than a direct computational proof with components.
      \end{enumerate}
  \end{problem}

  \begin{problem}
    \leavevmode\begin{enumerate}[(a)]
        \item Let $\bv=(2,6,3)$ and $\bw=(1,1,1)$. Calculate $\mathbf P_\bw \bv$ and $\mathbf P_\bw \mathbf P_\bw \bv$. What do you notice?
        \item Generalize part (a): Prove for all vectors $\bv,\bw$ with $\bw\neq 0$, that $\mathbf P_\bw(\mathbf P_\bw \bv)=\mathbf P_\bw \bv$, in other words projecting twice does the same thing as projecting once.
    \end{enumerate}
  \end{problem}

\subsection*{Challenge problem}
Find the set of all vectors which make a $\pi/3$ angle with the vector $(0,0,1)$. (If the angle was $\pi/2$ instead, we know that this would form the $xy$ plane.) Can you prove or disprove whether this set is a plane? What does it look like? What is an equation for this set?\footnote{Spoiler: It looks like a cone.}

\subsection*{Challenge problem}
How would you disprove that the four points $(1,5,3)$, $(2,-2,1)$, $(1,0,0)$, and $(0,0,0)$ lie in the same plane?

\end{document}
