\documentclass[11pt,oneside]{amsart}
\input{../preamble}

\title{MATH2202 Spring 2024\\
Discussion 7}
\date{March 12, 2024}

\theoremstyle{definition}
\newtheorem{question}{Question}

\begin{document}
  \maketitle

  The TA will demonstrate 1(a) and 1(b). You will need Mathematica for Problem 2.

  \begin{problem}
    The multivariable chain rule gives a way to prove a formula for implicit differentiation you may have seen before in AP calculus: Given the equation $F(x,y)=0$, then
    \[\frac{dy}{dx}=-\frac{F_x(x,y)}{F_y(x,y)}\]
    \begin{enumerate}[(a)]
      \item Prove this formula using the multivariable chain rule.
      \item Try it out yourself on the example $x^3y^2-6xy-1=0$. Compare with normal implicit differentiation.
    \end{enumerate}
  \end{problem}

  \begin{problem}\leavevmode
    \begin{enumerate}[(a)]
      \item Recall from class that $(0,0)$, $(18,0)$, and $(0,18)$ were all saddle points while $(6,6)$ was a local maximum for the function
      \[f(b,\ell)=b\ell(18-b-\ell).\]
      Plot this function in Mathematica in 3D and check that those points have the nature as stated. Also observe how the shape of the plot looks while you're at it.
      \item Also plot level curves in Mathematica to give another way to see how it looks. Use \texttt{ContourPlot}. Set the \texttt{Contours} option to a larger number such as 50. Can you see the local maximum as well as the 3 saddle points on this plot?
      \item Try finding and classifying critical points for the slightly different function
      \[f(b,\ell)=b\ell(18-2b-\ell).\]
      Then check with Mathematica using \texttt{ContourPlot}.
    \end{enumerate}
  \end{problem}

  \begin{problem}
    Consider the ellipse $x^2+2y^2=1$. With the help of the gradient vector, find a normal vector to the curve at $P=\left( \frac1{\sqrt2},\frac12 \right)$.
  \end{problem}

  \begin{problem}
    What does the Hessian test (another term for second derivative test) say about the function $f(x,y)=x^3+y^3$ at $(0,0)$? Show that in every direction through $(0,0)$ the second derivative is 0, meaning we have an inflection point in every direction.

    Another good example is $x^3+y^2-2y$ at $(0,1)$. Make a contour plot for this in Mathematica! In which directions do we have a local min/max/inflection point?
  \end{problem}
\end{document}
