\documentclass[11pt,oneside]{amsart}
\input{../preamble}

\title{MATH2202 Spring 2024\\
Discussion 12}
\date{April 23, 2024}

\theoremstyle{definition}
\newtheorem{question}{Question}

\begin{document}
  \maketitle

  The TA will demonstrate 1.

  \begin{problem}
    This is the example I promised yesterday. Let $\BF(x,y)=\frac{(-y,x)}{x^2+y^2}$.
    \begin{enumerate}[(a)]
      \item Sketch $\BF$.
      \item Show that $\nabla\times \BF=0$.
      \item Show that $\int_C\BF\cdot d\br\neq 0$, where $C$ is the unit circle starting and ending at $(1,0)$.
      \item This seems to contradict Theorem 5 on p.1079 (this is the curl test, stated equivalently). Explain which condition was not met and why.
    \end{enumerate}
  \end{problem}

  \begin{problem}
    This problem will be about line integrals over closed curves. All curves in this question are traversed counterclockwise.
    \begin{enumerate}[(a)]
      \item Let $\BF(x,y)=(1,x)$. Calculate $\int_C \BF\cdot d\br$ where $C$ is the unit circle starting and ending at $(1,0)$.
      \item Calculate $\int_C \BF\cdot d\br$ where $C$ is the unit circle starting and ending at $(0,1)$. (Hint: $(\sin(t),-\cos(t))$ for $0\leq t\leq 2\pi$ can parametrize this curve.)
      \item Calculate $\int_C \BF\cdot d\br$ where $C$ is the unit circle starting and ending at $(-1,0)$.
      \item You should've gotten the same answer for all 3 parts. Make a conjecture about line integrals over closed curves to explain why.
      \item Challenge: Prove this conjecture. You can do it by splitting the curve into two parts and rearranging.
    \end{enumerate}
  \end{problem}

  \begin{problem}
    Prove the second part of the grad-curl-div theorem from yesterday: $\nabla\cdot(\nabla\times\BF)=0$ for all twice continuously differentiable vector fields $\BF$. Start by writing $\BF$ as $(M,N,R)$.
  \end{problem}
\end{document}
