\documentclass[11pt,oneside]{amsart}
\input{../preamble}

\title{MATH2202 Spring 2024\\
Discussion 9}
\date{March 26, 2024}

\theoremstyle{definition}
\newtheorem{question}{Question}

\begin{document}
  \maketitle

  The TA will demonstrate 1(a) and 3(a).

  \begin{problem}\leavevmode
    \begin{enumerate}[(a)]
      \item Write down an equation that describes a cone with base radius $b$ and height $r$ sitting on top of the $xy$-plane.
      \item Calculate the volume of this cone by integration.
    \end{enumerate}
  \end{problem}

  \begin{problem}
    The graph of $r=\cos(2\theta)$, $0\leq\theta\leq 2\pi$ is a 4-leaf clover.
    \begin{enumerate}[(a)]
      \item Sketch it.
      \item Find the area of one of the leaves.

      (Hint: First, figure out the $\theta$ range that represents one of the leaves.)
    \end{enumerate}
  \end{problem}

  \begin{problem}\leavevmode
    \begin{enumerate}[(a)]
      \item Rotation by an angle $\theta$ around the origin in the $xy$-plane is a change of coordinates. Write down the functions $u(x,y)$ and $v(x,y)$ so that, for any given input $(x,y)$, the output $(u(x,y),v(x,y))$ represents the vector obtained from $(x,y)$ after rotating counterclockwise by the angle $\theta$ around the origin.
      \item Show that the Jacobian determinant of $(u(x,y),v(x,y))$ is 1.
      \item What does that indicate? Convince yourself it means that rotation preserves areas. Further convince yourself that this fact can also be arrived at by using the fact that rotation is an isometry, meaning it preserves distances and lengths.
    \end{enumerate}
  \end{problem}

\end{document}
